Stack Safety as a Security Property

### 2020-07-02 main.tex
More comments

\bcp{
Comments on the example:
\begin{itemize}
\item We can simplify a bit by renaming $f$ to $\mathit{main}$ and
relocating its stack to location 1000.  And maybe move its instructions to
start at 0.  (We can comment that on a real RISC-V this is where interrupt
handlers go, but POPL readers will not care too much.)  We should streamline
the example as much as possible.
\item Do we need the ``...''s in $f$?
%
\item We should also emphasize, again, that there is no heap (if there were,
we might put it above the stack or between the instructions and the stack).
%
\item We will also need to be careful about fonts ($f$ vs {\texttt f}),
capitalization of register names (sp vs SP), etc. Maybe typeset instructions
in small caps?  Maybe typeset everything that is part of a concrete machine
in small caps??
%
\item The colors are helpful in making the big page of instruction traces easier
on the eye; maybe they would be even more helpful if
corresponding colors appeared elsewhere in the example (on the instructions
themselves, e.g.).  But we need to be careful to keep it readable -- maybe
they should be background colors, like you've done, rather than text colors.
\item The first column of numbers is redundant with the second column (PC).
Suggest deleting the first.
\item Asking readers to wade through all this is asking quite a bit!  We
need to make sure we are going to get sufficient value out of the example
later (I think we can, but we should really convince ourselves) to make it
worthwhile.
\item Use ``='' instead of ``:'' in the ``summaries'' below the pictures.
And put a bit more space between the components.
\item
\end{itemize}
}

@@

\newcommand*{\tracerow}[5]{#1: & #2 & #3 & #4 & \text{#5}\\}

@@

    \tracerow{2}{104}{\sw ~ \rsp ~ r_0 ~ -8}{[1048] \leftarrow 0}{Store 0 as argument}
    \tracerow{3}{108}{\sw ~ \rsp ~ r_0 ~ -4}{[1052] \leftarrow 0}{Zero out result}

@@

    \tracerow{6}{144}{\sw ~ \rsp ~ \rra ~ -4}{[1056] \leftarrow 116}{Store return address}

@@

    \tracerow{9}{156}{\sw ~ \rsp ~ r_4 ~ -8}{[1060] \leftarrow 17}{Store 17 as the argument}
    \tracerow{10}{160}{\sw ~ \rsp ~ r_0 ~ -4}{[1064] \leftarrow 0 }{Zero out return}

@@

  \tracerow{13}{204}{\sw ~ \rsp ~ \rra ~ -4}{[1068] \leftarrow 168}{Store return address}
  \tracerow{14}{208}{\lw ~ r_4 ~ \rsp ~ -12}{r_4 \leftarrow 17}{Load the argument of $h$ to $r_4$}

@@

  \tracerow{16}{216}{\sw ~ \rsp ~ r_4 ~ -8}{[1064] \leftarrow 18}{Store $r_4$ as the return value}
  \tracerow{17}{220}{\lw ~ \rra ~ \rsp ~ -4}{\rra \leftarrow 168}{Load return address}
  \tracerow{18}{224}{\addi ~ \rsp ~ \rsp ~ -4}{\rsp \leftarrow 1068}{Restore the stack pointer of $g$}

@@

\begin{figure}
% After Return from H
% Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{49em}{1.8em}{$\downarrow$}
\MemoryLabel{48.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
\memory{3}{gray}
~$\cdots$
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{0}
\MemoryLabel{-20.6em}{0.6em}{116}
\MemoryLabel{-18.4em}{0.6em}{17}
\MemoryLabel{-16.2em}{0.6em}{18}
\\
\end{center}
% Register File
\[
\begin{array}{cccccc}
  PC: 168 & r_0 : 0 & \rra : 168 & \rsp : 1068 & r_4 : 18 & r_5 : ?
\end{array}
\]
% H - Body
\vspace*{0.2em}
\[
\begin{array}{l|l|l|l|l}
  \tracerow{20}{168}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
  \tracerow{21}{172}{\lw ~ r_5 ~ \rsp ~ -20}{r_5 \leftarrow 0}{Load the argument of $g$ to $r_5$}
  \tracerow{22}{176}{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow 18 + 0}{Add $r_4$ and $r_5$}
  \tracerow{23}{180}{\sw ~ \rsp ~ r_4 ~ -8}{[1056] \leftarrow 18}{Store $r_4$ as ther return of $g$}
  \tracerow{24}{184}{\addi ~ \rsp ~ \rsp ~ -8}{\rsp \leftarrow 1060}{Deallocate local state}
  \tracerow{25}{188}{\lw ~ \rra ~ \rsp ~ -4}{\rra \leftarrow 116}{Load return address}
  \tracerow{26}{192}{\addi ~ \rsp ~ \rsp ~ -4}{\rsp \leftarrow 1056}{Restore the stack pointer of $f$}
  \tracerow{27}{196}{\jalr ~ \rra ~ \rra ~ 0}{PC \leftarrow 116}{Return to $f$}
\end{array}
\]
% At Call to G
% Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{41em}{1.8em}{$\downarrow$}
\MemoryLabel{40.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg + result}}]%
\memory{6}{gray}
~$\cdots$
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{18}
\\
\end{center}
% Register File
\[
\begin{array}{cccccc}
  PC: 116 & r_0 : 0 & \rra : 116 & \rsp : 1056 & r_4 : 18 & r_5 : 0
\end{array}
\]
\vspace*{0.2em}
\[
\begin{array}{l|l|l|l|l}
  \tracerow{28}{116}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
  \tracerow{29}{120}{\sw ~ \rsp ~ r_4 ~ -12}{[1048] \leftarrow 18}{Set $y$ to the return value}
  \tracerow{30}{124}{\addi ~ \rsp ~ \rsp ~ -8}{\rsp \leftarrow 1048}{Deallocate Call Frame}
\end{array}
\]
\vspace*{0.2em}
% Ending Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
\MemoryLabel{35em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{8}{gray}
~$\cdots$
\MemoryLabel{-29.6em}{0.6em}{42}
\MemoryLabel{-27.4em}{0.6em}{0}
\\
\end{center}
% Starting Register File
\[
\begin{array}{cccccc}
  PC: 128 & r_0 : 0 & \rra : 116 & \rsp : 1048 & r_4 : 18 & r_5 : 0
\end{array}
\]
\caption{Detailed Execution Trace}
\label{fig:running-trace-b}
\end{figure}

@@

Formally a code map is a partial function from addresses to a function
identifier drawn from the set \(\FUNIDS\).
  %
\[\codemap \in \CODEMAPS = \WORDS \rightharpoonup \FUNIDS\]
%
In our running example, instructions 100-124 correspond to $f$,
instructions 140-196 correspond to $g$, and instructions 200-228
correspond to $h$.

@@

 \[\mathit{Cof}(\mach,\funid,n)(\component) =
  \begin{cases}
    (\LOWSEC,\HIGHINT) & \text{if } \component \in \WORDS
                         \text{ and for some } \funid, \codemap(\component) = \funid \\
    (\HIGHSEC,\HIGHINT) & \text{else if } \component \in \WORDS
                          \text{ and } \component \leq \mach(\SP) - n \\
    (\HIGHSEC,\LOWINT) & \text{else if } \component \in \WORDS
                         \text{ and } \component \geq \mach(\SP) \\
    (\LOWSEC,\LOWINT) & \text{else} \\
  \end{cases}\]

>>>>

\bcp{
Comments on the example:
\begin{itemize}
\item We can simplify a bit by renaming $f$ to $\mathit{main}$ and
relocating its stack to location 1000.  And maybe move its instructions to
start at 0.  (We can comment that on a real RISC-V this is where interrupt
handlers go, but POPL readers will not care too much.)  We should streamline
the example as much as possible.
\item Do we need the ``...''s in $f$?
%
\item We should also emphasize, again, that there is no heap (if there were,
we might put it above the stack or between the instructions and the stack).
%
\item We will also need to be careful about fonts ($f$ vs {\texttt f}),
capitalization of register names (sp vs SP), etc. Maybe typeset instructions
in small caps?  Maybe typeset everything that is part of a concrete machine
in small caps??
%
\item The colors are helpful in making the big page of instruction traces easier
on the eye; maybe they would be even more helpful if
corresponding colors appeared elsewhere in the example (on the instructions
themselves, e.g.).  But we need to be careful to keep it readable -- maybe
they should be background colors, like you've done, rather than text colors.
\item The first column of numbers is redundant with the second column (PC).
Suggest deleting the first.
\item Asking readers to wade through all this is asking quite a bit!  We
need to make sure we are going to get sufficient value out of the example
later (I think we can, but we should really convince ourselves) to make it
worthwhile.
\item Use ``='' instead of ``:'' in the ``summaries'' below the pictures.
And put a bit more space between the components.
\item Create a new column that just literally pronounces each instruction in
English, so people don't have to read RISC-V opcodes.
\item As discussed, we should move this to before section 2, we should
extend it with a bit more at the end so we can talk about the lazy policy
with the same example, we should add
in some bad variants, and we should consider discussing the attacker model
here (we need to discuss it someplace in any case!)
\end{itemize}
}

@@

\newcommand*{\tracerow}[5]{#1: & #2 & #3 & #4 & \text{#5}\\}
\newcommand{\negate}{\textrm{-}}

@@

    \tracerow{2}{104}{\sw ~ \rsp ~ r_0 ~ \negate 8}{[1048] \leftarrow 0}{Store 0 as argument}
    \tracerow{3}{108}{\sw ~ \rsp ~ r_0 ~ \negate 4}{[1052] \leftarrow 0}{Zero out result}

@@

    \tracerow{6}{144}{\sw ~ \rsp ~ \rra ~ \negate 4}{[1056] \leftarrow 116}{Store return address}

@@

    \tracerow{9}{156}{\sw ~ \rsp ~ r_4 ~ \negate 8}{[1060] \leftarrow 17}{Store 17 as the argument}
    \tracerow{10}{160}{\sw ~ \rsp ~ r_0 ~ \negate 4}{[1064] \leftarrow 0 }{Zero out return}

@@

  \tracerow{13}{204}{\sw ~ \rsp ~ \rra ~ \negate 4}{[1068] \leftarrow 168}{Store return address}
  \tracerow{14}{208}{\lw ~ r_4 ~ \rsp ~ \negate 12}{r_4 \leftarrow 17}{Load the argument of $h$ to $r_4$}

@@

  \tracerow{16}{216}{\sw ~ \rsp ~ r_4 ~ \negate 8}{[1064] \leftarrow 18}{Store $r_4$ as the return value}
  \tracerow{17}{220}{\lw ~ \rra ~ \rsp ~ \negate 4}{\rra \leftarrow 168}{Load return address}
  \tracerow{18}{224}{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow 1068}{Restore the stack pointer of $g$}

@@

\begin{figure}
% After Return from H
% Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{49em}{1.8em}{$\downarrow$}
\MemoryLabel{48.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
\memory{3}{gray}
~$\cdots$
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{0}
\MemoryLabel{-20.6em}{0.6em}{116}
\MemoryLabel{-18.4em}{0.6em}{17}
\MemoryLabel{-16.2em}{0.6em}{18}
\\
\end{center}
% Register File
\[
\begin{array}{cccccc}
  PC: 168 & r_0 : 0 & \rra : 168 & \rsp : 1068 & r_4 : 18 & r_5 : ?
\end{array}
\]
% H - Body
\vspace*{0.2em}
\[
\begin{array}{l|l|l|l|l}
  \tracerow{20}{168}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
  \tracerow{21}{172}{\lw ~ r_5 ~ \rsp ~ \negate 20}{r_5 \leftarrow 0}{Load the argument of $g$ to $r_5$}
  \tracerow{22}{176}{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow 18 + 0}{Add $r_4$ and $r_5$}
  \tracerow{23}{180}{\sw ~ \rsp ~ r_4 ~ \negate 8}{[1056] \leftarrow 18}{Store $r_4$ as ther return of $g$}
  \tracerow{24}{184}{\addi ~ \rsp ~ \rsp ~ \negate 8}{\rsp \leftarrow 1060}{Deallocate local state}
  \tracerow{25}{188}{\lw ~ \rra ~ \rsp ~ \negate 4}{\rra \leftarrow 116}{Load return address}
  \tracerow{26}{192}{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow 1056}{Restore the stack pointer of $f$}
  \tracerow{27}{196}{\jalr ~ \rra ~ \rra ~ 0}{PC \leftarrow 116}{Return to $f$}
\end{array}
\]
% At Call to G
% Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{41em}{1.8em}{$\downarrow$}
\MemoryLabel{40.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg + result}}]%
\memory{6}{gray}
~$\cdots$
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{18}
\\
\end{center}
% Register File
\[
\begin{array}{cccccc}
  PC: 116 & r_0 : 0 & \rra : 116 & \rsp : 1056 & r_4 : 18 & r_5 : 0
\end{array}
\]
\vspace*{0.2em}
\[
\begin{array}{l|l|l|l|l}
  \tracerow{28}{116}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
  \tracerow{29}{120}{\sw ~ \rsp ~ r_4 ~ \negate 12}{[1048] \leftarrow 18}{Set $y$ to the return value}
  \tracerow{30}{124}{\addi ~ \rsp ~ \rsp ~ \negate 8}{\rsp \leftarrow 1048}{Deallocate Call Frame}
\end{array}
\]
\vspace*{0.2em}
% Ending Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
\MemoryLabel{35em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{8}{gray}
~$\cdots$
\MemoryLabel{-29.6em}{0.6em}{42}
\MemoryLabel{-27.4em}{0.6em}{0}
\\
\end{center}
% Starting Register File
\[
\begin{array}{cccccc}
  PC: 128 & r_0 : 0 & \rra : 116 & \rsp : 1048 & r_4 : 18 & r_5 : 0
\end{array}
\]
\caption{Detailed Execution Trace}
\label{fig:running-trace-b}
\end{figure}

@@

Formally a code map\bcp{I'd love to change ``cdm'' to ``km''.  Should we
  also call it a ``kode map'' in English?  Maybe that's a bit too far.} is a
partial function from addresses to a function
identifier drawn from the set \(\FUNIDS\).
  %
\[\codemap \in \CODEMAPS = \WORDS \rightharpoonup \FUNIDS\]
%
In our running example, instructions 100-124 correspond to $f$,
instructions 140-196 correspond to $g$, and instructions 200-228
correspond to $h$.

@@

 \[\mathit{Cof}(\mach,\funid,n)(\component) =
  \begin{cases}
    (\LOWSEC,\HIGHINT) & \text{if } \component \in \WORDS
                         \text{ and for some } \funid, \codemap(\component) = \funid \\
    (\HIGHSEC,\HIGHINT) & \text{else if } \component \in \WORDS
                          \text{ and } \component \leq \mach(\SP) \negate  n \\
    (\HIGHSEC,\LOWINT) & \text{else if } \component \in \WORDS
                         \text{ and } \component \geq \mach(\SP) \\
    (\LOWSEC,\LOWINT) & \text{else} \\
  \end{cases}\]
